On-the-fly Algorithms for GKAT Equivalences \thanks{Identify applicable funding agency here. If none, delete this.}

### 2025-01-16 main.tex
check point recursive construction

The original algorithm for deciding GKAT equivalences~\cite{smolka_GuardedKleeneAlgebra_2020} requires the entire coalgebra to be known prior to the execution of the bisimulation algorithm; specifically, it is necessary to iterate through the entire coalgebra in order to identify the liveness of every single states, in order to perform the normalization operation.
This limitation poses challenges to design an efficient on-the-fly algorithm for GKAT.
In order to make the decision procedure scalable, we will need to merge the normalization and bisimulation procedure, so that our algorithm can normalize the coalgebra only when we need to.

In this section, we introduce an algorithm that merges bisimulation and normalization, where we only test the liveness of states when a disparity is noticed by the bisimulation algorithm.
For example, when deciding the trace equivalence between \(s ∈ S\) and \(u ∈ U\), if \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\), we will proceed to recurse on \(s'\) and \(u'\), without checking the liveness of \(s'\) and \(u'\).
However, if \(s\) rejects \(α\) instead of transitioning to \(s'\), we will then check whether \(u'\) is dead.

We present several sound and complete conditions for finite-trace equivalence between states of GKAT coalgebras in~\Cref{thm:recursive-construction}, which also serves as the core correctness theorem for our equivalence-checking algorithm i.e.~\Cref{alg:bisim}.
This algorithm mostly consists of recursively checking the conditions in~\Cref{thm:recursive-construction} with minor optimizations, which we will outline later.

\begin{definition}[greedy bisimulation]\label{def:norm-bisim}
    Given two GKAT coalgebra \(S\) and \(U\), and a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \({≈} ⊆ S × U\) is a GKAT coalgebra that contains \(∼\) with the following transition function:
    \[
        δ_≈((s, u), α) ≜ \begin{cases}
            \accept & \begin{aligned}
                & \text{if } δ_S(s, α) = \accept \\[-3px]
                & \text{and } δ_U(s, α) = \accept
            \end{aligned}\\[5px]
            ((s', u'), p) & 
                \begin{aligned}
                    & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
                    & \text{and } u \transvia{α ∣ p}_{U} u' 
                \end{aligned}\\[5px]
            \reject & \text{otherwise}
        \end{cases}
    \]
    We call \(∼\) \emph{the underlying bisimulation} of \(≈\).
\end{definition}

\begin{lemma}\label{thm:dead-construction-greedy-bisim}
    Given any eager bisimulation \({≈} ⊆ S × U\) with the underlying bisimulation \({∼}\), the following relation \({∼_d} ⊆ S × U\):
    \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:norm-bisim}, which implies \({≈} ⊑ {∼_d}\).
    Furthermore, \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
\end{lemma}

\begin{proof}
    We will need to show \(∼_d\) is closed under the transition of \(δ_{≈}\) in~\Cref{def:norm-bisim}.
    Specifically, if \(s ∼_d u\) and \(s \transvia{α ∣ p}_{S} s'\) and \(u \transvia{α ∣ p} u'\), then \(s' ∼_d u'\).
    By case analysis on the liveness of \(s'\) and \(u'\) in \(S\) and \(U\):
    \begin{itemize}
        \item If both \(s'\) and \(u'\) are dead, then \(s ∼_d u\) by definition of \(∼_d\).
        \item If both \(s'\) and \(u'\) are live, then \(s, s'\) and \(u, u'\) are live in \(S\) and \(U\) respectively. 
        Thus, \(s \transvia{α ∣ p}_{\norm(S)} s'\), and \(u \transvia{α ∣ p}_{\norm(U)} u'\).
        Therefore, \(s' ∼ u'\) by definition of bisimulation on \(\norm(S)\) and \(\norm(U)\); and \(s ∼_d u'\).
        \item If \(s'\) is live but \(u'\) is dead, then \(s\) is live, and \(s ∼ u\).
        However, \(s \transvia{α ∣ p}_{\norm(S)} s'\) yet \(u \transRej{\norm(U)}{u}\), which violates the condition for bisimulation, thus this case cannot happen.
        \item If \(u'\) is live but \(s'\) is dead, this case also cannot happen similar to the previous case.
    \end{itemize}

@@

\begin{lemma}\label{thm:norm-bisim-exists-unique}
    For any bisimulation between two normalized GKAT coalgebra \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
\end{lemma}

@@

\begin{lemma}\label{thm:norm-bisim-not-bisim-then-component-dead}
    Every state in \((s, u) ∈ {≈} ⊆ S × U\) but not in \({∼} ⊆ \norm(S) × \norm(U)\) then \(s\) is dead in \(S\), \(u\) is dead in \(U\).
\end{lemma}

@@

\begin{theorem}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
    Given a pair of states in a greedy bisimulation \(s ≈ u\) where \({≈} ⊆ S × U\), then the following are equivalent
    \begin{itemize}
        \item \((s, u)\) is live (or dead) in \({≈}\);
        \item \(s\) is live (or dead) in \(S\);  
        \item \(u\) is live (or dead) in \(U\).
    \end{itemize}
    Furthermore, if \((s, u)\) is in the underlying bisimulation \(∼\) of \(≈\), then all the above is also equivalent to \((s, u)\) is live (or dead) in \(∼\).
\end{theorem}

@@

\begin{lemma}\label{thm:greedy-bisim-iff-norm-bisim}
    Given two GKAT coalgebras \(S\) and \(U\), and a relational GKAT coalgebra \({≈} ⊆ S × U\) with the transition \(δ_≈\) as in~\cref{def:norm-bisim}.
    \(≈\) is a greedy bisimulation if and only if \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\)
\end{lemma}

\begin{proof}
    The \(⟹\) direction, we let \(∼_d\) be defined as in~\Cref{thm:dead-construction-greedy-bisim}.
    Because \({≈} ⊑ {∼_d}\) and~\Cref{thm:norm-sub-coalg}, \(\norm(≈) ⊑ \norm(∼_d)\). 
    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), then we can show that \(\norm(≈)\) is a bisimulation by factoring through \(\norm(∼_d)\), in the category of GKAT coalgebras:
    \begin{align*}
        & \norm(≈) 
        \xrightarrow{i} \norm(∼_d) 
        \xrightarrow{π₁} \norm(S) \\
        & \norm(≈) 
        \xrightarrow{i} \norm(∼_d) 
        \xrightarrow{π₂} \norm(U) 
    \end{align*}
    because inclusion homomorphism \(i\) do not change its input, the two homomorphism obtained by bisimulation is indeed \(π₁: \norm(≈) → \norm(S)\) and \(π₂: \norm(≈) → \norm(U)\), implying both projections of \(\norm(≈)\) are homomorphism, meaning \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).

@@

@@

\begin{proof}
    \(⟹\) direction, because normalization perserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
    And because \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), \(⟦s⟧_S = ⟦u⟧_U\).

@@

@@

    The condition~\labelcref{itm:transition-bisim} holds because when \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), then \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) by definition of greedy bisimulation (\Cref{def:norm-bisim}).

    The condition~\labelcref{itm:transition-dead} holds, let \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', q)\) and \(p ≠ q\). 
    \begin{itemize}
        \item If \(s ∼ u\), then we can show condition~\labelcref{itm:transition-dead} by case analysis on the liveness of \(s'\) and \(u'\):
        if \(s', u'\) are not both dead, then the condition of \({∼} ⊆ \norm(S) × \norm(U)\) cannot be satisfied, hence \((s', u') ∉ {∼}\), which by~\Cref{thm:norm-bisim-not-bisim-then-component-dead} means that \(s'\) and \(u'\) are both dead, contradicting the premise.
        Therefore, \(s'\) and \(u'\) needs to be both dead.
        \item If \(s\) and \(u\) are both dead, then by~\Cref{thm:dead-iff-all-reachable-dead}, both \(s\) and \(u\) can only transition to dead state, hence \(s'\) and \(u'\) are both dead.
    \end{itemize}

@@

    For all pair of states \((s, u) ∈ \norm(≈)\) and any atom \(α ∈ \At\), we will verify all the conditions of a homomorphism.
    For the accepting case, by the definition of \(δ_{≈}\):
    \begin{align*}
        (s, u) \transAcc{\norm(≈)}{α}
        & \text{ iff } (s, u) \transAcc{≈}{α} \\  
        & \text{ implies } s \transAcc{≈}{α}
    \end{align*}
    For the transition case, because \((s, u)\) is live in \(≈\) implies that \(s\) is live in \(S\) (\Cref{thm:greedy-bisim-perserve-liveness}):
    \begin{align*}
        & (s, u) \transvia{α ∣ p}_{\norm(S)} (s', u') \\
        & \text{iff }
            (s, u) \transvia{α ∣ p}_{S} (s', u')
            \text{ and \(s, u\) is live in \(≈\)} \\  
        & \text{implies }
            s \transvia{α ∣ p}_S s' 
            \text{ and \(s\) is live in \(S\)} \\  
        & \text{implies }
            s \transvia{α ∣ p}_{\norm(S)} s' 
    \end{align*}
    Finally, the rejection case: \((s, u) \transRej{\norm(≈)}{α}\) either when \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) and \((s', u')\) is dead, or \((s, u) \transRej{≈}{α}\).
    We will show that both cases will lead to \(s \transRej{\norm(S)}{α}\).
    The \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) case can be proven by the definition of \(δ_≈\) and~\Cref{thm:greedy-bisim-perserve-liveness}:
    \begin{align*}
        & (s, u) \transvia{α ∣ p}_{≈} (s', u') 
            \text{ and \((s', u')\) is dead in \(≈\)} \\
        \text{implies } & s \transvia{α ∣ p}_{S} s' 
            \text{ and \(s'\) is dead in \(S\)} \\  
        \text{implies } & s \transRej{\norm(S)}{α}.
    \end{align*}
    Then the hardest case is when \((s, u) \transRej{≈}{α}\) because by definition of \(δ_≈\) (\Cref{def:norm-bisim}) there are many cases that can lead to rejection.
    We will need to eliminate irrelevant cases using the condition provided by the construction.
    We proceed by case analysis on dynamics of \(s\) and \(u\) on \(α\); first listing some cases that contradicts the premise \((s, u) \transRej{≈}{α}\):
    \begin{itemize}
        \item If \(s\) accepts \(α\) in \(S\), then \(u\) necessarily accepts \(α\) in \(U\) by condition~\labelcref{itm:acc-condition}, which implies \((s, u) \transAcc{≈}{α}\), contradicting the premise.
        \item Similarly, if \(u\) accepts \(α\) in \(U\), then \((s, u) \transAcc{≈}{α}\), contradicting the premise.
        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\) then \((s, u) \transvia{α ∣ p}_≈ (s', u;)\), contradicting the premise.
    \end{itemize}
    Then we list the other cases, which indeed satisfy \((s, u) \transRej{≈}{α}\), and show that all of them leads to \(s \transRej{\norm(S)}{α}\), using the condition in the theorem:
    \begin{itemize}
        \item If \(s \transRej{S}{α}\), then \(s \transRej{\norm(S)}{α}\).
        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transRej{U}{α}\), then by condition~\labelcref{itm:rej-or-dead}, \(s'\) is dead, and \(s \transRej{\norm(S)}{α}\).
        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ q}_U u'\), then by condition~\labelcref{itm:transition-dead}, both \(s'\) and \(u'\) are dead, and \(s \transRej{\norm(S)}{α}\).
        \qedhere
    \end{itemize}
\end{proof}

@@

\begin{proof}
    First, recall that by~\Cref{thm:greedy-bisim-iff-norm-bisim} \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
    Because \(S ⊑ S + U\) and \(U ⊑ S + U\) and normalization preserves sub-coalgebra~\Cref{thm:norm-sub-coalg}, we have \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\); and by~\Cref{thm:sub-coalg-preserve-liveness}, \(s ∈ S\) is dead in \(S + U\) if and only if it is dead in \(S\), and similarly for \(u ∈ U\).

@@

    Let \(≊\) be the greedy bisimulation generated by \(≡\), then \(≊\) has the same carrier as \(≡\).
    All the \((s, u) ∈ {≊}\) are either \(s ≡ u\) or \(s\) and \(u\) are dead in \(S\) and \(U\), which implies that both are dead in \(S + U\).
    However, \(≡\) already contains all the dead pairs (\cref{thm:bisim-between-dead}), therefore for all \((s, u) ∈ {≊}\), they are also in \(≡\), implying \({≡} ⊇ {≊}\).
    Finally, by definition of greedy bisimulation (\Cref{def:norm-bisim}), \({≡} ⊆ {≊}\).
    Therefore, \(≡\) and \(≊\) share the same carrier and \(≊\) is a greedy bisimulation that is also an equivalence relation, which also contains \((s, u)\):
    \[(s, u) ∈ {≈} = \norm(≈) ⊆ {≡} ⊆ {≊},\]
    where all the \(=\) and \(⊆\) are orders on the carrier set, instead of the coalgebra.

    The \(⟸\) direction can be split into two cases, by whether \((s, u)\) is in the underlying bisimulation of \(≊\), which we denote as \(≃\): 
    \begin{itemize}
        \item If \(s ≃ u\), then we can reflect \(≃\)  through \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) (\Cref{thm:sub-coalg-preserve-bisim}) to obtain a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
        Taking the greedy bisimulation over \(∼\) will yield a desired greedy bisimulation \(≈\) s.t. \(s ≈ u\).
        \item If \((s, u)\) are not in \(≃\), then they are both dead in \(S + U\) (\Cref{thm:norm-bisim-not-bisim-then-component-dead}), which implies \(s\) and \(u\) are respectively dead in \(S\) and \(U\) (\Cref{thm:sub-coalg-preserve-liveness}).
        Then we take \(≈\) to be greedy bisimulation generated form the maximal bisimulation between \(S\) and \(U\), which contains \(s\) and \(u\) (\Cref{thm:bisim-between-dead}).
        \qedhere
    \end{itemize}
\end{proof}

% % \begin{definition}[Normalizable Bisimulation]
% %     A relation \({≃} ⊆ S × U\) is a normalizable bisimulation if it has a GKAT coalgebra structure \(δ_≃\), and \(\norm(≃)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
% % \end{definition}

% % \begin{lemma}[Preservation of Liveness]
% %     For any normalizable bisimulation \({≃} ⊆ S × U\) if \(s ≃ u\), then \(s\) is live in \(S\) if and only if \(u\) is live in \(U\).
% % \end{lemma}

% % \begin{proof}
% %     Because \(\norm(≃)\) has two projection homomorphism to \(\norm(S)\) and \(\norm(U)\), then 
% % \end{proof}

% % \begin{theorem}[Correctness]
% %     If there exists a normalizable bisimulation \({≃} ⊆ S × U\) s.t. \(s ∼ u\) for \(s ∈ S\) and \(u ∈ U\), then \(⟦s⟧_S = ⟦u⟧_U\).
% % \end{theorem}

% % \begin{theorem}[Construction]
% %     Given two states \(s ∈ S, u ∈ U\) then there exists a normalizable bisimulation \({≃'} ⊆ S × U\) if and only if there exists a relation \({≃} ⊆ S × U\) that satisfy the following conditions:
% %     \begin{enumerate}
% %         \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
% %         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
% %         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
% %         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
% %     \end{enumerate}
% % \end{theorem}

% % \begin{proof}
% %     The \(⟸\) direction can be shown that 
% % \end{proof}

% % \begin{definition}[greedy bisimulation]\label{def:norm-bism}
% %     A relation \({≃} ⊆ S × U\) on two GKAT coalgebra \(S\) and \(U\) is a greedy bisimulation if and only if for any two states \(s ≃ u\), all the following holds:
% %     \begin{enumerate}
% %         \item\label{itm:acc-condition} for all \(α ∈ \At\), \(δ_{S}(s, α) = \accept ⟺ δ_{U}(u, α) = \accept\);
% %         \item\label{itm:rej-or-dead} if \(s\) reject \(α\) but \(u\) transitions to \(u'\), then \(u'\) is dead; similarly when \(u\) rejects \(α\);  
% %         \item\label{itm:transition-bisim} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', p)\), then \(s' ≃ u'\).
% %         \item\label{itm:transition-dead} If \(δ_{S}(s, α) = (s', p)\) and \(δ_{U}(u, α) = (u', q)\), s.t. \(p ≠ q\), then both \(s'\) and \(t'\) are dead.
% %     \end{enumerate}
% %     And \(≃\) can be endowed with the following GKAT coalgebra structure:
% %     \[
% %         δ_≃((s, u), α) ≜ \begin{cases}
% %             \accept & \begin{aligned}
% %                 & \text{if } δ_S(s, α) = \accept \\[-3px]
% %                 & \text{and } δ_U(s, α) = \accept
% %             \end{aligned}\\[5px]
% %             ((s', u'), p) & 
% %                 \begin{aligned}
% %                     & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
% %                     & \text{and } u \transvia{α ∣ p}_{U} u' 
% %                 \end{aligned}\\[5px]
% %             \reject & \text{otherwise}
% %         \end{cases}
% %     \]
% % \end{definition}

>>>>

The original algorithm for deciding GKAT equivalences~\cite{smolka_GuardedKleeneAlgebra_2020} requires the entire coalgebra to be known prior to the execution of the bisimulation algorithm; specifically, it is necessary to iterate through the entire coalgebra in order to identify the liveness of every single states and perform the normalization operation.
This limitation poses challenges to design an efficient on-the-fly algorithm for GKAT.
To make the equivalence-checking procedure scalable, we propose to greedily run the bisimulation algorithm and only invoke liveness detection when a discrepancy is found. 
For example, when deciding the trace equivalence between \(s ∈ S\) and \(u ∈ U\), if \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', p)\), we will proceed to recurse on \(s'\) and \(u'\), without checking the liveness of \(s'\) and \(u'\).
However, if \(s\) rejects \(α\) instead of transitioning to \(s'\), we will then check whether \(u'\) is dead.

This approach offers several benefit: first, because the liveness detection is only called when necessary, thus it is unlikely to iterate through the entire coalgebra; second, the algorithm can short-circuit on a counter-example, allowing termination even before all the states are explored; finally, the algorithm supports on-the-fly constructions like derivatives, because the coalgebra no longer is required to be fully constructed.

The correctness argument of this on-the-fly algorithm is centered around the following coalgebraic structure, which we named ``greedy bisimulation''.

\begin{definition}[Greedy Bisimulation]\label{def:greedy-bisim}
    Given two GKAT coalgebra \(S\) and \(U\), and a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \({≈} ⊆ S × U\) is the least GKAT coalgebra that contains \(∼\) with the following transition function:
    \[
        δ_≈((s, u), α) ≜ \begin{cases}
            \accept & \begin{aligned}
                & \text{if } δ_S(s, α) = \accept \\[-3px]
                & \text{and } δ_U(s, α) = \accept
            \end{aligned}\\[5px]
            ((s', u'), p) & 
                \begin{aligned}
                    & \text{if } s \transvia{α ∣ p}_{S} s' \\[-3px]
                    & \text{and } u \transvia{α ∣ p}_{U} u' 
                \end{aligned}\\[5px]
            \reject & \text{otherwise}
        \end{cases}
    \]
    We call \(∼\) \emph{the underlying bisimulation} of \(≈\).
\end{definition}

Given any bisimulation \({∼} ⊆ S × U\), we can construct a GKAT coalgebra \(∼_d\) s.t. it is almost a greedy bisimulation, i.e. it is closed under \(δ_≈\) in~\Cref{def:greedy-bisim}, and contains \(∼\), except it is not the \emph{least} GKAT coalgebra that satisfy these properties.

\begin{lemma}\label{thm:dead-construction-greedy-bisim}
    Given any eager bisimulation \({≈} ⊆ S × U\) with the underlying bisimulation \({∼}\), the following relation \({∼_d} ⊆ S × U\):
    \[{∼_d} ≜ {∼} ∪ \{(s, u) ∣ \text{\(s\) and \(u\) are dead in \(S\) and \(U\)}\},\]
    form a GKAT coalgebra the transition function \(δ_≈\) in~\Cref{def:greedy-bisim}, which implies \({≈} ⊑ {∼_d}\).
    Furthermore, \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
\end{lemma}

\begin{proof}
    We will need to show \(∼_d\) is closed under the transition of \(δ_{≈}\) in~\Cref{def:greedy-bisim}.
    Specifically, if \(s ∼_d u\) and \(s \transvia{α ∣ p}_{S} s'\) and \(u \transvia{α ∣ p} u'\), then \(s' ∼_d u'\).
    By case analysis on the liveness of \(s'\) and \(u'\) in \(S\) and \(U\):
    \begin{itemize}
        \item If both \(s'\) and \(u'\) are dead, then \(s ∼_d u\) by definition of \(∼_d\).
        \item If both \(s'\) and \(u'\) are live, then \(s, s'\) and \(u, u'\) are live in \(S\) and \(U\) respectively. 
        Thus, \(s \transvia{α ∣ p}_{\norm(S)} s'\), and \(u \transvia{α ∣ p}_{\norm(U)} u'\).
        Therefore, \(s' ∼ u'\) by definition of bisimulation on \(\norm(S)\) and \(\norm(U)\); and \(s ∼_d u'\).
        \item If \(s'\) is live but \(u'\) is dead, then \(s\) is live, and \(s ∼ u\).
        However, \(s \transvia{α ∣ p}_{\norm(S)} s'\) yet \(u \transRej{\norm(U)}{u}\), which violates the condition for bisimulation, thus this case cannot happen.
        \item If \(u'\) is live but \(s'\) is dead, this case also cannot happen similar to the previous case.
    \end{itemize}

@@

Although the proof of~\Cref{thm:dead-construction-greedy-bisim} is basically by unfolding the definition, \(∼_d\) turns out to be a important structure. 
Specifically, \(∼_d\) serves a connection between \(≈\) and its underlying bisimulation \(∼\), and the order \({∼} ⊆ {≈} ⊑ {~_d}\) will reveal many structural properties of \(≈\).
We show the following corollary to demonstrate the usefulness of \(∼_d\).

\begin{corollary}\label{thm:norm-bisim-exists-unique}
    For any bisimulation between two normalized GKAT coalgebra \({∼} ⊆ \norm(S) × \norm(U)\), a greedy bisimulation \(≈\) with \(∼\) as underlying bisimulation exists and is unique.
\end{corollary}

@@

\begin{corollary}\label{thm:greedy-bisim-dead-or-bisim}
    For every pair of states \((s, u)\) in a greedy bisimulation \({≈} ⊆ S × U\), either \((s, u)\) is in the underlying bisimulation \(∼\), or both \(s\) and \(u\) are dead in \(S\) and \(U\).
\end{corollary}

@@

\begin{corollary}[Perservation Of Liveness]\label{thm:greedy-bisim-perserve-liveness}
    Given a pair of states in a greedy bisimulation \(s ≈ u\) where \({≈} ⊆ S × U\), then the following are equivalent
    \begin{itemize}
        \item \((s, u)\) is live (or dead) in \({≈}\);
        \item \(s\) is live (or dead) in \(S\);  
        \item \(u\) is live (or dead) in \(U\).
    \end{itemize}
    Furthermore, if \((s, u)\) is in the underlying bisimulation \(∼\) of \(≈\), then all of above are also equivalent to \((s, u)\) is live (or dead) in \(∼\).
\end{corollary}

@@

\begin{corollary}\label{thm:greedy-bisim-iff-norm-bisim}
    Given two GKAT coalgebras \(S\) and \(U\), and a relational GKAT coalgebra \({≈} ⊆ S × U\) with the transition \(δ_≈\) as in~\cref{def:greedy-bisim}.
    \(≈\) is a greedy bisimulation if and only if \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\)
\end{corollary}

\begin{proof}
    The \(⟹\) direction, we let \(∼_d\) be defined as in~\Cref{thm:dead-construction-greedy-bisim}.
    Because \({≈} ⊑ {∼_d}\) and~\Cref{thm:norm-sub-coalg}, \(\norm(≈) ⊑ \norm(∼_d)\). 
    Because \(\norm(∼_d)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\), therefore \(\norm(≈)\) is a bisimulation by~\cref{thm:}

@@

We then establish the correctness of greedy bisimulation, i.e. it is equivalent to trace equivalence.

@@

\begin{proof}
    \(⟹\) direction, because normalization preserves carrier set, therefore \((s, u) ∈ \norm(≈)\).
    And because \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\) (\Cref{thm:greedy-bisim-iff-norm-bisim}), \(⟦s⟧_S = ⟦u⟧_U\).

@@

Despite proving many desirable structural properties, the definition of greedy bisimulation is still coalgebraic.
To design an equivalence-checking algorithm to construct a greedy bisimulation still requires turning the coalgebra into a set of concrete criteria, that can be checked by an algorithm.
In the following theorem, we will present four conditions, and the algorithm will be largely based on recursively checking these conditions to establish a greedy bisimulation.

@@

    The condition~\labelcref{itm:transition-bisim} holds because when \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\), then \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) by definition of greedy bisimulation (\Cref{def:greedy-bisim}).

    The condition~\labelcref{itm:transition-dead} holds, let \(δ_S(s, α) = (s', p)\) and \(δ_U(u, α) = (u', q)\) and \(p ≠ q\). 
    \begin{itemize}
        \item If \(s ∼ u\), then we can show condition~\labelcref{itm:transition-dead} by case analysis on the liveness of \(s'\) and \(u'\):
        if \(s', u'\) are not both dead, then the condition of \({∼} ⊆ \norm(S) × \norm(U)\) cannot be satisfied, hence \((s', u') ∉ {∼}\), which by~\Cref{thm:greedy-bisim-dead-or-bisim} means that \(s'\) and \(u'\) are both dead, contradicting the premise.
        Therefore, \(s'\) and \(u'\) needs to be both dead.
        \item If \(s\) and \(u\) are both dead, then by~\Cref{thm:dead-iff-all-reachable-dead}, both \(s\) and \(u\) can only transition to dead state, hence \(s'\) and \(u'\) are both dead.
    \end{itemize}

@@

    For all pair of states \((s, u) ∈ \norm(≈)\) and any atom \(α ∈ \At\), we will verify all the conditions of a homomorphism.
    For the accepting case, by the definition of \(δ_{≈}\):
    \begin{align*}
        (s, u) \transAcc{\norm(≈)}{α}
        & \text{ iff } (s, u) \transAcc{≈}{α} \\  
        & \text{ implies } s \transAcc{≈}{α}
    \end{align*}
    For the transition case, because \((s, u)\) is live in \(≈\) implies that \(s\) is live in \(S\) (\Cref{thm:greedy-bisim-perserve-liveness}):
    \begin{align*}
        & (s, u) \transvia{α ∣ p}_{\norm(S)} (s', u') \\
        & \text{iff }
            (s, u) \transvia{α ∣ p}_{S} (s', u')
            \text{ and \(s, u\) is live in \(≈\)} \\  
        & \text{implies }
            s \transvia{α ∣ p}_S s' 
            \text{ and \(s\) is live in \(S\)} \\  
        & \text{implies }
            s \transvia{α ∣ p}_{\norm(S)} s' 
    \end{align*}
    Finally, the rejection case: \((s, u) \transRej{\norm(≈)}{α}\) either when \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) and \((s', u')\) is dead, or \((s, u) \transRej{≈}{α}\).
    We will show that both cases will lead to \(s \transRej{\norm(S)}{α}\).
    The \((s, u) \transvia{α ∣ p}_{≈} (s', u')\) case can be proven by the definition of \(δ_≈\) and~\Cref{thm:greedy-bisim-perserve-liveness}:
    \begin{align*}
        & (s, u) \transvia{α ∣ p}_{≈} (s', u') 
            \text{ and \((s', u')\) is dead in \(≈\)} \\
        \text{implies } & s \transvia{α ∣ p}_{S} s' 
            \text{ and \(s'\) is dead in \(S\)} \\  
        \text{implies } & s \transRej{\norm(S)}{α}.
    \end{align*}
    Then the hardest case is when \((s, u) \transRej{≈}{α}\) because by definition of \(δ_≈\) (\Cref{def:greedy-bisim}) there are many cases that can lead to rejection.
    We will need to eliminate irrelevant cases using the condition provided by the construction.
    We proceed by case analysis on dynamics of \(s\) and \(u\) on \(α\); first listing some cases that contradicts the premise \((s, u) \transRej{≈}{α}\):
    \begin{itemize}
        \item If \(s\) accepts \(α\) in \(S\), then \(u\) necessarily accepts \(α\) in \(U\) by condition~\labelcref{itm:acc-condition}, which implies \((s, u) \transAcc{≈}{α}\), contradicting the premise.
        \item Similarly, if \(u\) accepts \(α\) in \(U\), then \((s, u) \transAcc{≈}{α}\), contradicting the premise.
        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ p}_U u'\) then \((s, u) \transvia{α ∣ p}_≈ (s', u;)\), contradicting the premise.
    \end{itemize}
    Then we list the other cases, which indeed satisfy \((s, u) \transRej{≈}{α}\), and show that all of them leads to \(s \transRej{\norm(S)}{α}\), using the condition in the theorem:
    \begin{itemize}
        \item If \(s \transRej{S}{α}\), then \(s \transRej{\norm(S)}{α}\).
        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transRej{U}{α}\), then by condition~\labelcref{itm:rej-or-dead}, \(s'\) is dead, and \(s \transRej{\norm(S)}{α}\).
        \item If \(s \transvia{α ∣ p}_S s'\) and \(u \transvia{α ∣ q}_U u'\), then by condition~\labelcref{itm:transition-dead}, both \(s'\) and \(u'\) are dead, and \(s \transRej{\norm(S)}{α}\).
        \qedhere
    \end{itemize}
\end{proof}

@@

\begin{proof}
    First, recall that by~\Cref{thm:greedy-bisim-iff-norm-bisim} \(\norm(≈)\) is a bisimulation between \(\norm(S)\) and \(\norm(U)\).
    Because \(S ⊑ S + U\) and \(U ⊑ S + U\) and normalization preserves sub-coalgebra (\Cref{thm:norm-sub-coalg}), we have \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\); and by~\Cref{thm:sub-coalg-preserve-liveness}, \(s ∈ S\) is dead in \(S + U\) if and only if it is dead in \(S\), and similarly for \(u ∈ U\).

@@

    Let \(≊\) be the greedy bisimulation generated by \(≡\), then \(≊\) has the same carrier as \(≡\).
    All the \((s, u) ∈ {≊}\) are either \(s ≡ u\) or \(s\) and \(u\) are dead in \(S\) and \(U\), which implies that both are dead in \(S + U\).
    However, \(≡\) already contains all the dead pairs (\cref{thm:bisim-between-dead}), therefore for all \((s, u) ∈ {≊}\), they are also in \(≡\), implying \({≡} ⊇ {≊}\).
    Finally, by definition of greedy bisimulation (\Cref{def:greedy-bisim}), \({≡} ⊆ {≊}\).
    Therefore, \(≡\) and \(≊\) share the same carrier and \(≊\) is a greedy bisimulation that is also an equivalence relation, which also contains \((s, u)\):
    \[(s, u) ∈ {≈} = \norm(≈) ⊆ {≡} ⊆ {≊},\]
    where all the \(=\) and \(⊆\) are orders on the carrier set, instead of the coalgebra.

    The \(⟸\) direction can be split into two cases, by whether \((s, u)\) is in the underlying bisimulation of \(≊\), which we denote as \(≃\): 
    \begin{itemize}
        \item If \(s ≃ u\), then we can reflect \(≃\)  through \(\norm(S) ⊑ \norm(S + U)\) and \(\norm(U) ⊑ \norm(S + U)\) (\Cref{thm:sub-coalg-preserve-bisim}) to obtain a bisimulation \({∼} ⊆ \norm(S) × \norm(U)\) and \(s ∼ u\).
        Taking the greedy bisimulation over \(∼\) will yield a desired greedy bisimulation \(≈\) s.t. \(s ≈ u\).
        \item If \((s, u)\) are not in \(≃\), then they are both dead in \(S + U\) (\Cref{thm:greedy-bisim-dead-or-bisim}), which implies \(s\) and \(u\) are respectively dead in \(S\) and \(U\) (\Cref{thm:sub-coalg-preserve-liveness}).
        Then we take \(≈\) to be greedy bisimulation generated form the maximal bisimulation between \(S\) and \(U\), which contains \(s\) and \(u\) (\Cref{thm:bisim-between-dead}).
        \qedhere
    \end{itemize}
\end{proof}
